---
title: "Talin Loop Paper Data Analysis"
author: "Jannik Buhr"
date: 2022-04-11
format:
  html:
    theme: [default, hits.scss]
    code-tools: true
    toc: true
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk: 
    fig.path: 'results/plots/'
    fig.align: 'center'
    fig.show: 'hold'
    dev: 'ragg_png'
    fig.ext: 'png'
    fig.retina: 3
    comment: ''
    out.width: '100%'
    R.options:
      tidyverse.quiet = true
editor: source

---

# Setup

```{r}
source("./R/globals.R")
source("./R/functions.R")
```

# Data processing pipeline

```{r tar-make}
#| eval: false
#| include: false
tar_make()
```

```{r tar-network}
tar_visnetwork(callr_arguments = list(show = FALSE))
```


## Notes

- Data loss due to disk failure:
  - rotational sampling
    - run2/angle_204: corrupted 00143.dat.csv, 00166.dat.csv
    - run3/angle_234: corrupted after first 4 frames
    - run4/angle_348: corrupted

# Plots

## Load data

```{r load-data}
tar_load(data_path_prefix)
tar_load(CUTOFF)
tar_load(from_first_contact)
tar_load(interacting)
tar_load(sequence)
f0f1_run1 <- tar_read(f0f1_rot_sample_distances_per_run, branches = 1:60)
tar_load(rolling)
tar_load(stuck)
tar_load(ferm_interacting)
tar_load(ferm_annotation)
tar_load(ferm_colors)
tar_load(bool_colors)
tar_load(tln1_ferm_lat_pull)
tar_load(f0f1_pulling_smooth)
tar_load(f0f1_pulling_peaks)
tar_load(ferm_pip_sites)
tar_load(f0f1_pulling_interacting)

sem <- function(x) sd(x) / sqrt(length(x))
```

## Rotational sampling of F0F1

```{r f0f1-ri-angle-npip}
paths <- c("{data_path_prefix}papers/paper-talin-loop/assets/blender/render-rotsample/frame0001.png",
    "{data_path_prefix}papers/paper-talin-loop/assets/blender/render-rotsample/frame0004.png",
    "{data_path_prefix}papers/paper-talin-loop/assets/blender/render-rotsample/frame0003.png",
    "{data_path_prefix}papers/paper-talin-loop/assets/blender/render-rotsample/frame0002.png",
    "{data_path_prefix}papers/paper-talin-loop/assets/blender/render-rotsample/frame0001.png") |>  map_chr(glue)

render_png <- function(p) {grid::rasterGrob(png::readPNG(p))}
pngs <- map(paths, render_png)

plt <- from_first_contact |>  
  group_by(i, angle) |>  
  summarise(
    n_pip = mean(n_pip)
  ) |>  
  ungroup() |>  
  complete(i = 1:197, angle = seq(0, 354, 6), fill = list(n_pip = 0)) |>  
  ggplot(aes(i, angle, fill = n_pip)) + 
  geom_raster() +
  scale_fill_gradient(low = "white", high = HITS_BLUE) +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  labs(
    x = "residue index",
    y = "angle [deg]",
    fill = expression(bar(n)[pip2])
  )

grobs <- ggplotGrob(plt)
plt <- plt +
  guides(fill = guide_colorbar(barheight = unit(0.7, "grobheight", data = grobs$grobs),
                               frame.colour = "black"))

anot <- ggplot(ferm_annotation[1:3,]) +
  aes(
    ymin = 0, ymax = 10,
    xmin = imin, xmax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_text(aes(y = 5, x = (imin + imax)/2, label = domain),
            color = "black") +
  theme_void() +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  scale_fill_discrete(type = ferm_colors)

vmd <- wrap_plots(pngs, ncol = 1)

main <- (plt / anot) +
  plot_layout(heights = c(20,1)) +
  plot_annotation()

((vmd / plot_spacer()) | main) +
  plot_layout(widths = c(1,10))
```

```{r f0f1-ri-npip}
plt <- from_first_contact |>  
  group_by(i) |> 
  summarise(n_pip = mean(n_pip)) |>  
  left_join(sequence, by = "i") |>  
  ggplot(aes(x = i, y = n_pip)) + 
  annotate(geom = "rect",
           xmin = 140, xmax = 164,
           ymin = 0, ymax = Inf,
           color = NA,
           fill = "black",
           alpha = 0.05) +
  geom_col(aes(fill = PI),
           alpha = 1,
           width = 0.8,
           show.legend = FALSE) +
  geom_text(aes(label = res, color = PI),
            fontface = "bold",
            vjust = -0.2,
            data = ~ filter(.x, n_pip > 0.1)) +
  scale_color_gradient2(low = HITS_MAGENTA,
                        mid = "grey90", high = HITS_BLUE, midpoint = 7) +
  scale_fill_gradient2(low = HITS_MAGENTA,
                       mid = "grey90", high = HITS_BLUE, midpoint = 7) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(y = expression(n[PIP[2]])) +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  labs(
    x = "residue index",
    y = expression(bar(n)[PIP[2]]),
    ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grobs <- ggplotGrob(plt)

plt <- plt +
  guides(color = guide_colorbar(barheight = unit(0.7, "grobheight", data = grobs$grobs),
                               frame.colour = "black"))


anot <- ggplot(ferm_annotation[1:3,]) +
  aes(
    ymin = 0, ymax = 10,
    xmin = imin, xmax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_text(aes(y = 5, x = (imin + imax)/2, label = domain),
            color = "black") +
  theme_void() +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  scale_fill_discrete(type = ferm_colors)

(plt / anot) +
  plot_layout(heights = c(20,1)) +
  plot_annotation()
```

```{r f0f1-angle-frame-npip}
plt <- interacting |>  
  group_by(frame, angle) |>  
  summarise(pip = mean(n_pip)) |>  
  ggplot(aes(angle, frame, fill = pip)) +
  geom_raster() +
  scale_fill_gradient(low = "white", high = HITS_BLUE) +
  guides(color = guide_colorbar(barheight = 16)) +
  labs(
    fill = expression(bar(n)[PIP[2]]),
    y = "time [ns]",
    x = "angle [deg]"
  )
  
grobs <- ggplotGrob(plt)

plt +
  guides(fill = guide_colorbar(barheight = unit(0.7, "grobheight", data = grobs$grobs),
                               frame.colour = "black"))
```

```{r f0f1-retention}
f0f1_stickiness <- rolling |>  
  group_by(run, angle) |>  
  slice_max(time) |>  
  summarise(stuck = n_i > 0) |>  
  ungroup()

f0f1_time_ri <- interacting |>  
  group_by(run, angle, frame) |>  
  summarise(n_i = sum(n_pip > 0)) |>  
  ungroup() |>  
  left_join(f0f1_stickiness)

plt1 <- f0f1_time_ri |>  
  filter(!is.na(stuck)) |>  
  mutate(stuck = fct_rev(factor(stuck))) |>  
  ggplot(aes(frame, n_i, group = paste(run, angle), color = stuck)) +
  geom_line(aes(alpha = stuck)) +
  geom_smooth(aes(group = stuck, color = stuck), se = FALSE,
              data = ~filter(.x, stuck == TRUE)) +
  scale_color_manual(values = bool_colors, name = "retention") +
  scale_alpha_manual(values = c(`TRUE` = 0.08, `FALSE` = 0.6)) +
  labs(x = "time [ns]",
       y = "# interacting residues") +
  coord_cartesian(clip = "off") +
  guides(color = "none", alpha = "none") +
  theme(
    plot.margin = unit(c(1, 0, 1, 1), "lines")
  )

plt2 <- stuck |>  
  ggplot(aes(x = 0, fill = stuck, n, label = paste0(n))) +
  geom_col(position = "stack") +
  geom_text(vjust = 1.1, position = "stack", fontface = "bold", size = 3, color = "white") +
  scale_x_continuous(breaks = NULL) +
  scale_fill_manual(values = c( "grey30", HITS_MAGENTA, HITS_BLUE)) +
  labs(x = "", y = "", title = "summary") +
  theme(
    legend.position = "top",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = unit(c(1, 1, 0, 0), "lines")
  ) +
  guides(fill = guide_legend(title = ""))


(plt1 | plt2) +
  plot_layout(widths = c(15, 1), guides = 'collect') &
  theme(legend.position = "bottom")
```

```{r f0f1-distance-cutoff}
plt <- f0f1_run1 |>  
  filter(r != 1) |>  
  ggplot(aes(r, y = stat(density))) +
  geom_density(fill = HITS_BLUE, alpha = 0.6) +
  geom_vline(xintercept = CUTOFF, lty = 2) +
  annotate("text", x = CUTOFF, y = 1,
           label = glue("r = {CUTOFF} nm"), hjust = -0.5,
           fontface = "bold") +
  labs(
    x = "r [nm]",
    y = "#",
    caption = "r ≥ 1 not shown due to cutoff"
  ) +
  coord_cartesian(xlim = c(0, 0.9))

plt
```


## Vertical pulling of F0F1

```{r f0f1-vert-pull}
plt_contacts <- f0f1_pulling_interacting |>  
  group_by(run, frame) |>  
  summarise(contacts = sum(n_pip > 0)) |>  
  ggplot(aes(frame * 0.03, contacts)) +
  geom_line(aes(group = run), alpha = 0.6) +
  geom_line(data = ~filter(.x, run == 4), aes(group = run), alpha = 0.6, color = HITS_MAGENTA) +
  labs(x = "distance [nm]",
       y = "# interacting residues"
       )

plt_curves <- f0f1_pulling_smooth |> 
  ggplot(aes(time / 1e3 *  0.03, f)) +
  geom_line(aes(group = run), alpha = 0.3) +
  geom_smooth(aes(group = run), se = FALSE, color = HITS_BLUE) +
  geom_smooth(data = ~filter(.x, run == 4), aes(group = run), se = FALSE, color = HITS_MAGENTA) +
  labs(x = "distance [nm]",
       y = "force [pN]",
       caption = "pulling rate: 0.03 nm/ns") +
  coord_cartesian(ylim = c(0, 150))

(plt_contacts / plt_curves)
```

```{r}
f0f1_pulling_interacting |> 
  group_by(run) |> 
  filter(frame == max(frame)) |> 
  summarise(sum(n_pip > 0))
```

- Run 4 looks interesting.
- PIP~2~s get pulled out of the membrane, one lodges itself near F0, two more get pulled out by the loop.

```{r f0f1-vert-pull-contacts}
# loop: 135 to 168
df <- f0f1_pulling_interacting |>
  arrange(run, frame, i) |>
  mutate(loop = between(i, 135, 168)) |>
  group_by(run, frame, loop) |>  
  summarise(contacts = sum(n_pip > 0)) |> 
  group_by(run, loop) |>
  mutate(contacts = zoo::rollmean(contacts, 15, NA)) |>
  pivot_wider(names_from = loop, values_from = contacts) |> 
  ungroup()

head_tail <- function(data, ...) {
  bind_rows(
    slice_max(data, ...),
    slice_min(data, ...)
  )
}

points <- df |>  
  filter(!is.na(`FALSE`), !is.na(`TRUE`)) |>  
  group_by(run) |>  
  slice_head()

arrows_end <- df |>  
  filter(!is.na(`FALSE`), !is.na(`TRUE`)) |>  
  group_by(run) |>  
  head_tail(frame, n = 2) |>  
  arrange(run, frame) |>  
  mutate(end = rep(1:(n() %/%  2), each = 2)) |>
  filter(end == 2)

df |>
  left_join(
    f0f1_pulling_smooth |> 
      select(run, time, f) |> 
      mutate(frame = time %/% 1e3) |> 
      group_by(run, frame) |> 
      summarise(f = mean(f))
  ) |> 
  ggplot(aes(`TRUE`, `FALSE`, color = f, group = paste(run))) +
  geom_path() +
  geom_point(data = points, color = "black") +
  geom_path(
    aes(group = paste(run, end)),
    color = "black",
    data = arrows_end,
    arrow = arrow(angle = 15, ends = "last", type = "closed",
                  length = unit(0.4, "lines")),
  ) +
  scale_color_viridis_c(direction = -1) +
  coord_equal() +
  labs(
    x = "loop contacts",
    y = "non loop contacts",
    color = "F [pN]"
  )
```

```{r f0f1-vert-pull-contacts-time}
df |> 
  ggplot(aes(`TRUE`, `FALSE`, color = frame , group = paste(run))) +
  geom_path() +
  geom_point(data = points, color = "black") +
  geom_path(
    aes(group = paste(run, end)),
    color = "black",
    data = arrows_end,
    arrow = arrow(angle = 15, ends = "last", type = "closed",
                  length = unit(0.4, "lines")),
  ) +
  scale_color_gradient(high = "#98cde9", low = "#003063") +
  coord_equal() +
  guides(color = "none") +
  labs(
    x = "loop contacts",
    y = "non loop contacts",
    color = ""
  )
```

```{r f0f1-vert-pull-residues}
# rate: 0.03 nm/ns
rate <- 0.03

plt <- f0f1_pulling_interacting |> 
  group_by(frame, i) |> 
  summarise(n_pip = mean(n_pip)) |> 
  mutate(d = frame * rate) |> 
  ggplot(aes(i, d, fill = n_pip)) +
  geom_raster() +
  # facet_wrap(~run) +
  scale_fill_gradient(low = "white", high = HITS_BLUE) +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  guides(fill = "none") +
  labs(y = "pulling distance [nm]",
       x = "residue index"
       # caption = "pulling rate: 0.03 nm/ns"/
       )

anot <- ggplot(ferm_annotation[1:3,]) +
  aes(
    ymin = 0, ymax = 10,
    xmin = imin, xmax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_text(aes(y = 5, x = (imin + imax)/2, label = domain),
            color = "black") +
  theme_void() +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  scale_fill_discrete(type = ferm_colors)

(plt / anot) +
  plot_layout(heights = c(20,1)) +
  plot_annotation()
```

## Equilibrium simulaions of FERM

```{r ferm-time-ri-npip-single-run}
plt <- ferm_interacting |> 
  filter(run == 6) |> 
  ggplot(aes(frame, i, fill = n_pip)) +
  geom_raster() +
  scale_fill_gradient(low = "white", high = HITS_BLUE, breaks = c(0, 1, 2, 3)) +
  guides(fill = guide_legend(show.limits = TRUE)) +
  labs(
    x = "time [ns]",
    y = "residue index",
    fill = expression(n[pip[2]])
  ) +
  theme(legend.position = "bottom", legend.key = element_rect(color = "black")) +
  lims(y = c(0, 400))

anot <- ggplot(ferm_annotation) +
  aes(
    xmin = 0, xmax = 10,
    ymin = imin, ymax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_hline(data = ferm_pip_sites, aes(yintercept = ri), color = "red") +
  geom_text(aes(x = 5, y = (imin + imax)/2, label = domain), color = "black") +
  theme_void() +
  scale_fill_discrete(type = ferm_colors)

(plt | anot) +
  plot_layout(widths = c(25, 1)) +
  plot_annotation()
```

```{r ferm-time-ri-npip-all}
ferm_interacting |> 
  filter(run != 7) |> 
  ggplot(aes(frame, i, fill = n_pip)) +
  geom_raster() +
  facet_wrap(~ run, labeller = label_both) +
  scale_fill_gradient(low = "white", high = HITS_BLUE, breaks = c(0, 1, 2, 3)) +
  guides(fill = guide_legend(show.limits = TRUE)) +
  labs(
    x = "time [ns]",
    y = "residue index",
    fill = expression(n[pip[2]])
  ) +
  theme(legend.position = "bottom",
        panel.background = element_rect(color = "black", fill = NA)) 
```

```{r ferm-ri-npip}
plt <- ferm_interacting |> 
  group_by(i) |>   
  summarise(n_pip = mean(n_pip)) |> 
  left_join(sequence) |> 
  ggplot(aes(x = i, y = n_pip)) + 
  annotate(geom = "rect",
           xmin = 140, xmax = 164,
           ymin = 0, ymax = Inf,
           color = NA,
           fill = "black",
           alpha = 0.05) +
  geom_col(aes(fill = PI),
           alpha = 1,
           width = 0.8,
           show.legend = FALSE) +
  geom_text(aes(label = res, color = PI),
            fontface = "bold",
            vjust = -0.2,
            data = ~ filter(.x, n_pip > 0.1)) +
  scale_color_gradient2(low = HITS_MAGENTA,
                        mid = "grey90", high = HITS_BLUE, midpoint = 7) +
  scale_fill_gradient2(low = HITS_MAGENTA,
                       mid = "grey90", high = HITS_BLUE, midpoint = 7) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(breaks = seq(0, 400, 50)) +
  labs(y = expression(n[PIP[2]])) +
  coord_cartesian(clip = "off", xlim = c(0, 400)) +
  labs(
    x = "residue index",
    y = expression(bar(n)[PIP[2]])
    ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grobs <- ggplotGrob(plt)

plt <- plt +
  guides(color = guide_colorbar(barheight = unit(0.7, "grobheight", data = grobs$grobs),
                               frame.colour = "black"))

anot <- ggplot(ferm_annotation) +
  aes(
    ymin = 0, ymax = 10,
    xmin = imin, xmax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_vline(data = ferm_pip_sites, aes(xintercept = ri), color = "red") +
  geom_text(aes(y = 5, x = (imin + imax)/2, label = domain),
            color = "black") +
  theme_void() +
  coord_cartesian(clip = "off", xlim = c(0, 400)) +
  scale_fill_discrete(type = ferm_colors)


(plt / anot) +
  plot_layout(heights = c(20,1)) +
  plot_annotation()
```
