---
title: "Talin Loop Paper Data Analysis"
author: "Jannik Buhr"
date: 2022-04-11
format:
  html:
    theme: [default, hits.scss]
    code-tools: true
    code-fold: true
    toc: true
execute:
  echo: true
  warning: false
  message: false
knitr:
  opts_chunk: 
    fig.path: 'results/plots/'
    fig.align: 'center'
    fig.show: 'hold'
    dev: 'ragg_png'
    fig.ext: 'png'
    fig.retina: 3
    comment: ''
    out.width: '100%'
    R.options:
      tidyverse.quiet = true
editor: source
---

:::{.callout-tip}
To reproduce the graphics generated here, download the raw data from the archive and replace the value of the target `data_path_prefix` in `_targets.R` with the path to that data.
Then use the next two code chunks to load packages and functions and run the targets pipeline (`tar_make()`).
The raw data will then be processed such that you can load it in this document.
:::

# Setup

```{r}
source("./R/globals.R")
```

# Data processing pipeline

```{r tar-make}
#| eval: false
#| include: false
tar_make()
```

```{r tar-network}
tar_visnetwork(callr_arguments = list(show = FALSE))
```

# Plots

Asp125 and Glu126 missing -> sequence shifted by 2 afterwards.

## Load data

```{r}
f0f1_run1 <- tar_read(f0f1_rot_sample_distances_per_run, branches = 1:60)
tar_load(CUTOFF)
tar_load(from_first_contact)
tar_load(interacting)
tar_load(sequence)
tar_load(rolling)
tar_load(stuck)
tar_load(ferm_interacting)
tar_load(ferm_annotation)
tar_load(ferm_colors)
tar_load(bool_colors)
tar_load(f0f1_pulling_smooth)
tar_load(f0f1_pulling_peaks)
tar_load(ferm_pip_sites)
tar_load(f0f1_pulling_interacting)
tar_load(ferm_rmsf)

aa <- read_tsv("./assets/tables/aa.tsv") |>
  mutate(
    charge = str_remove(`Net charge at pH 7.4`, ",.+")
  ) |>
  select(`1`, `3`, charge)

CHARGE_COLORS = c(
  "Negative" = HITS_MAGENTA,
  "Positive" = HITS_BLUE,
  "Neutral" = "grey50"
)
```

## Rotational sampling of F0F1

- Data loss due to disk failure:
  - run4/angle_348: corrupted
  - run2/angle_204: corrupted only frames 00143 and 00166
  - run3/angle_234: corrupted after first 4 frames

```{r}
anti_join(
  expand_grid(run = 1:6, angle = seq(0, 354, 6)),
  distinct(interacting, run, angle)
)
```

```{r}
interacting |> 
  drop_na() |> 
  count(run, angle) |> 
  filter(n != max(n))
```

```{r}
anti_join(
  expand_grid(frame = 0:200, i = 1:197),
  interacting |> 
    filter(run == 2, angle == 204) |> 
    distinct(frame, i)
  )
```

```{r}
interacting |> 
  drop_na() |> 
  group_by(run, angle) |> 
  summarise(frame = max(frame)) |> 
  ungroup() |> 
  filter(frame != max(frame))
```

```{r}
#| label: f0f1-ri-angle-npip
paths <- c("./assets/blender/render-rotsample/frame0001.png",
    "./assets/blender/render-rotsample/frame0004.png",
    "./assets/blender/render-rotsample/frame0003.png",
    "./assets/blender/render-rotsample/frame0002.png",
    "./assets/blender/render-rotsample/frame0001.png") |>map_chr(glue)

render_png <- function(p) grid::rasterGrob(png::readPNG(p))
pngs <- map(paths, render_png)

plt <- from_first_contact |>
  group_by(i, angle) |>
  summarise(
      n_pip = mean(n_pip)
      ) |>
  ungroup() |>
  complete(i = 1:197, angle = seq(0, 354, 6), fill = list(n_pip = 0)) |>
  ggplot(aes(i, angle, fill = n_pip)) + 
  geom_raster() +
  scale_fill_gradient(low = "white", high = HITS_BLUE) +
  scale_y_continuous(breaks = seq(0, 360, 90)) +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  labs(
      x = "residue index",
      y = "angle [deg]",
      fill = expression(bar(n)[pip2])
      )

grobs <- ggplotGrob(plt)
plt <- plt +
guides(
    fill = guide_colorbar(
      barheight = unit(0.7, "grobheight", data = grobs$grobs),
      frame.colour = "black")
)

anot <- ggplot(ferm_annotation[1:3, ]) +
  aes(
    ymin = 0, ymax = 10,
    xmin = imin, xmax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_text(aes(y = 5, x = (imin + imax) / 2, label = domain),
            color = "black") +
  theme_void() +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  scale_fill_discrete(type = ferm_colors)

vmd <- wrap_plots(pngs, ncol = 1)

main <- (plt / anot) +
  plot_layout(heights = c(20, 1)) +
  plot_annotation()

((vmd / plot_spacer()) | main) +
  plot_layout(widths = c(1, 10))
```

```{r}
update <- function(data, row, colname, value) {
  col <- rlang::as_name(rlang::enquo(colname))
  data[row, col] <- value
  data
}
```

```{r f0f1-ri-npip}
plt <- from_first_contact |>
  group_by(i) |> 
  summarise(n_pip = mean(n_pip)) |>
  left_join(sequence, by = "i") |>
  left_join(aa, by = c('res' = '1')) |>
  update(1, charge, "Positive") |>
  ggplot(aes(x = i, y = n_pip)) + 
  annotate(geom = "rect",
           xmin = 140, xmax = 164,
           ymin = 0, ymax = Inf,
           color = NA,
           fill = "black",
           alpha = 0.05) +
  geom_col(aes(fill = charge),
           alpha = 1,
           width = 0.8,
           show.legend = TRUE) +
  geom_text(aes(label = res, color = charge),
            fontface = "bold",
            vjust = -0.2,
            data = ~ filter(.x, n_pip > 0.1), show.legend = FALSE) +
  scale_fill_manual(
    values = CHARGE_COLORS, aesthetics = c("fill", "color"),
    name = "Net Charge at pH 7.4"
  ) +
  scale_y_continuous(breaks = seq(0, 360, 90), expand = expansion(mult = c(0, 0.1))) +
  labs(y = expression(n[PIP[2]])) +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  labs(
    x = "residue index",
    y = expression(bar(n)[PIP[2]]),
    ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = c(0, 1),
    legend.justification = c(0, 1),
    legend.title = element_text(size = 12)
  )

grobs <- ggplotGrob(plt)

anot <- ggplot(ferm_annotation[1:3, ]) +
  aes(
    ymin = 0, ymax = 10,
    xmin = imin, xmax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_text(aes(y = 5, x = (imin + imax) / 2, label = domain),
            color = "black") +
  theme_void() +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  scale_fill_discrete(type = ferm_colors)

(plt / anot) +
  plot_layout(heights = c(20, 1)) +
  plot_annotation()
```

```{r f0f1-angle-frame-npip}
plt <- interacting |>
  group_by(frame, angle) |>
  summarise(pip = mean(n_pip)) |>
  ggplot(aes(angle, frame, fill = pip)) +
  geom_raster() +
  scale_fill_gradient(low = "white", high = HITS_BLUE) +
  scale_x_continuous(breaks = seq(0, 360, 90)) +
  guides(color = guide_colorbar(barheight = 16)) +
  labs(
    fill = expression(bar(n)[PIP[2]]),
    y = "time [ns]",
    x = "angle [deg]"
  )
  
grobs <- ggplotGrob(plt)

plt + guides(fill = guide_colorbar(barheight = unit(0.7, "grobheight",
  data = grobs$grobs),
  frame.colour = "black"))
```

```{r}
#| label: si-f0f1-first-contact-frame
interacting |>
  group_by(run, angle, i) |>
  arrange(frame) |>
  filter(n_pip > 0) |>
  slice_min(frame) |>
  group_by(i, angle) |>
  summarise(frame = mean(frame)) |>
  ungroup() |>
  mutate(
    region = case_when(
     between(i, 135, 168) ~ " ",
     between(i, 1, 84) ~ "f0",
     between(i, 85, 207) ~ "f1",
     TRUE ~ "rest"
    ),
  ) |>
  ggplot(aes(region, frame)) +
  geom_violin(aes(fill = region)) +
  # ggbeeswarm::geom_quasirandom(width = 0.4) +
  stat_summary(fun = "mean", geom = "crossbar", width = 0.7, size = 1) +
  scale_x_discrete(labels = c("f1-loop", "f0", "f1"), name = "") +
  scale_fill_discrete(type = ferm_colors) +
  guides(fill = "none") +
  labs(y = "first contact time [ns]")
```

```{r}
interacting |>
  group_by(run, angle, i) |>
  arrange(frame) |>
  filter(n_pip > 0) |>
  slice_min(frame) |>
  group_by(i, angle) |>
  summarise(frame = mean(frame)) |>
  ungroup() |>
  mutate(
    region = case_when(
     between(i, 135, 168) ~ "f1-loop",
     between(i, 1, 84) ~ "f0",
     between(i, 85, 207) ~ "f1",
     TRUE ~ "rest"
    ),
  ) |>
  (\(x) t.test(frame ~ region == "f1-loop", data = x))()
```

```{r}
#| label: si-f0f1-angle-first-to-find
interacting |>
  group_by(run, angle, i) |>
  arrange(frame) |>
  filter(n_pip > 0) |>
  slice_min(frame) |>
  ungroup() |>
  drop_na() |>
  mutate(
    region = case_when(
     between(i, 135, 168) ~ " ",
     between(i, 1, 84) ~ "f0",
     between(i, 85, 207) ~ "f1",
     TRUE ~ "rest"
    ),
    angle = cut(angle, breaks = seq(0, 360, by = 60), include.lowest = TRUE)
  ) |>
  ggplot(aes(angle, fill = region)) +
  geom_bar(position = "dodge2") +
  scale_fill_discrete(type = ferm_colors, labels = c("f1-loop", "f0", "f1")) +
  scale_y_continuous(expand = c(0, 1)) +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) +
  labs(y = "# region made first contact")
```

```{r f0f1-retention}
f0f1_stickiness <- rolling |>
  group_by(run, angle) |>
  slice_max(time) |>
  summarise(stuck = n_i > 0) |>
  ungroup()

f0f1_time_ri <- interacting |>
  group_by(run, angle, frame) |>
  summarise(n_i = sum(n_pip > 0)) |>
  ungroup() |>
  left_join(f0f1_stickiness)

plt1 <- f0f1_time_ri |>
  filter(!is.na(stuck)) |>
  mutate(stuck = fct_rev(factor(stuck))) |>
  ggplot(aes(frame, n_i, group = paste(run, angle), color = stuck)) +
  geom_line(aes(alpha = stuck)) +
  geom_smooth(aes(group = stuck, color = stuck), se = FALSE,
              data = ~filter(.x, stuck == TRUE)) +
  scale_color_manual(values = bool_colors, name = "retention") +
  scale_alpha_manual(values = c(`TRUE` = 0.08, `FALSE` = 0.6)) +
  labs(x = "time [ns]",
       y = "# interacting residues") +
  coord_cartesian(clip = "off") +
  guides(color = "none", alpha = "none") +
  theme(
    plot.margin = unit(c(1, 0, 1, 1), "lines")
  )

plt2 <- stuck |>
  ggplot(aes(x = 0, fill = stuck, n, label = paste0(n))) +
  geom_col(position = "stack") +
  geom_text(
    vjust = 1.1,
    position = "stack",
    fontface = "bold",
    size = 3,
    color = "white"
  ) +
  scale_x_continuous(breaks = NULL) +
  scale_fill_manual(values = c("grey30", HITS_MAGENTA, HITS_BLUE)) +
  labs(x = "total", y = "", title = "") +
  theme(
    legend.position = "top",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = unit(c(1, 1, 0, 0), "lines")
  ) +
  guides(fill = guide_legend(title = ""))


(plt1 | plt2) +
  plot_layout(widths = c(15, 1), guides = 'collect') &
  theme(legend.position = "bottom")
```

```{r f0f1-distance-cutoff}
plt <- f0f1_run1 |>
  filter(r != 1) |>
  ggplot(aes(r, y = stat(density))) +
  geom_density(fill = HITS_BLUE, alpha = 0.6) +
  geom_vline(xintercept = CUTOFF, lty = 2) +
  annotate("text", x = CUTOFF, y = 1,
           label = glue("r = {CUTOFF} nm"), hjust = -0.5,
           fontface = "bold") +
  labs(
    x = "r [nm]",
    y = "#",
    caption = "r ≥ 1 not shown due to cutoff"
  ) +
  coord_cartesian(xlim = c(0, 0.9))

plt
```

## Vertical pulling of F0F1

```{r f0f1-vert-pull}
plt_contacts <- f0f1_pulling_interacting |>
  group_by(run, frame) |>
  summarise(contacts = sum(n_pip > 0)) |>
  ggplot(aes(frame * 0.03, contacts)) +
  geom_line(aes(group = run, lty = (run == 4)), alpha = 0.6) +
  labs(
    x = "distance [nm]",
    y = "# interacting residues"
  ) +
  guides(lty = "none")

plt_curves <- f0f1_pulling_smooth |> 
  ggplot(aes(time / 1e3 *  0.03, f)) +
  geom_line(aes(group = run, lty = (run == 4)), alpha = 0.3) +
  geom_smooth(aes(group = run, lty = (run == 4)), se = FALSE, color = HITS_BLUE) +
  labs(x = "distance [nm]",
       y = "force [pN]",
       caption = "pulling rate: 0.03 nm/ns") +
  coord_cartesian(ylim = c(0, 150)) +
  guides(lty = "none")

(plt_contacts / plt_curves)
```

```{r}
f0f1_pulling_interacting |> 
  group_by(run) |> 
  filter(frame == max(frame)) |> 
  summarise(sum(n_pip > 0))
```

-   Run 4 looks interesting.
-   PIP~2~s get pulled out of the membrane, one lodges itself near F0, two more get pulled out by the loop.

```{r f0f1-vert-pull-contacts}
# loop: 135 to 168
df <- f0f1_pulling_interacting |>
  arrange(run, frame, i) |>
  mutate(loop = between(i, 135, 168)) |>
  group_by(run, frame, loop) |>
  summarise(contacts = sum(n_pip > 0)) |> 
  group_by(run, loop) |>
  mutate(contacts = zoo::rollmean(contacts, 15, NA)) |>
  pivot_wider(names_from = loop, values_from = contacts) |> 
  ungroup()

head_tail <- function(data, ...) {
  bind_rows(
    slice_max(data, ...),
    slice_min(data, ...)
  )
}

points <- df |>
  filter(!is.na(`FALSE`), !is.na(`TRUE`)) |>
  group_by(run) |>
  slice_head()

arrows_end <- df |>
  filter(!is.na(`FALSE`), !is.na(`TRUE`)) |>
  group_by(run) |>
  head_tail(frame, n = 2) |>
  arrange(run, frame) |>
  mutate(end = rep(1:(n() %/%  2), each = 2)) |>
  filter(end == 2)

df |>
  left_join(
    f0f1_pulling_smooth |> 
      select(run, time, f) |> 
      mutate(frame = time %/% 1e3) |> 
      group_by(run, frame) |> 
      summarise(f = mean(f))
  ) |> 
  ggplot(aes(`TRUE`, `FALSE`, color = f, group = paste(run))) +
  geom_path() +
  geom_point(data = points, color = "black") +
  geom_path(
    aes(group = paste(run, end)),
    color = "black",
    data = arrows_end,
    arrow = arrow(angle = 15, ends = "last", type = "closed",
                  length = unit(0.4, "lines")),
  ) +
  scale_color_viridis_c(direction = -1) +
  coord_equal() +
  labs(
    x = "loop contacts",
    y = "non loop contacts",
    color = "F [pN]"
  )
```

```{r f0f1-vert-pull-contacts-time}
df |> 
  ggplot(aes(`TRUE`, `FALSE`, color = frame, group = paste(run))) +
  geom_path() +
  geom_point(data = points, color = "black") +
  geom_path(
    aes(group = paste(run, end)),
    color = "black",
    data = arrows_end,
    arrow = arrow(angle = 15, ends = "last", type = "closed",
                  length = unit(0.4, "lines")),
  ) +
  scale_color_gradient(high = "#98cde9", low = "#003063") +
  coord_equal() +
  guides(color = "none") +
  labs(
    x = "loop contacts",
    y = "non loop contacts",
    color = ""
  )
```

```{r f0f1-vert-pull-residues}
# rate: 0.03 nm/ns
rate <- 0.03

plt <- f0f1_pulling_interacting |> 
  group_by(frame, i) |> 
  summarise(n_pip = mean(n_pip)) |> 
  mutate(d = frame * rate) |> 
  ggplot(aes(i, d, fill = n_pip)) +
  geom_raster() +
  # facet_wrap(~run) +
  scale_fill_gradient(low = "white", high = HITS_BLUE) +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  guides(fill = "none") +
  labs(y = "pulling distance [nm]",
       x = "residue index"
       # caption = "pulling rate: 0.03 nm/ns"/
       )

anot <- ggplot(ferm_annotation[1:3, ]) +
  aes(
    ymin = 0, ymax = 10,
    xmin = imin, xmax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_text(aes(y = 5, x = (imin + imax) / 2, label = domain),
            color = "black") +
  theme_void() +
  coord_cartesian(clip = "off", xlim = c(0, 207)) +
  scale_fill_discrete(type = ferm_colors)

(plt / anot) +
  plot_layout(heights = c(20, 1)) +
  plot_annotation()
```

## Equilibrium simulations of FERM

```{r ferm-time-ri-npip-single-run}
plt <- ferm_interacting |> 
  filter(run == 6) |> 
  ggplot(aes(frame, i, fill = n_pip)) +
  geom_raster() +
  scale_fill_gradient(low = "white", high = HITS_BLUE, breaks = c(0, 1, 2, 3)) +
  guides(fill = guide_legend(show.limits = TRUE)) +
  labs(
    x = "time [ns]",
    y = "residue index",
    fill = expression(n[pip[2]])
  ) +
  theme(
    legend.position = "bottom", legend.key = element_rect(color = "black")
  ) +
  lims(y = c(0, 400))

anot <- ggplot(ferm_annotation) +
  aes(
    xmin = 0, xmax = 10,
    ymin = imin, ymax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_hline(data = ferm_pip_sites, aes(yintercept = ri), color = "red") +
  geom_text(
      aes(x = 5, y = (imin + imax) / 2, label = domain),
      color = "black"
  ) +
  theme_void() +
  scale_fill_discrete(type = ferm_colors)

(plt | anot) +
  plot_layout(widths = c(25, 1)) +
  plot_annotation()
```

```{r ferm-time-ri-npip-all}
ferm_interacting |> 
  filter(run != 7) |> 
  ggplot(aes(frame, i, fill = n_pip)) +
  geom_raster() +
  facet_wrap(~ run, labeller = label_both) +
  scale_fill_gradient(low = "white", high = HITS_BLUE, breaks = c(0, 1, 2, 3)) +
  guides(fill = guide_legend(show.limits = TRUE)) +
  labs(
    x = "time [ns]",
    y = "residue index",
    fill = expression(n[pip[2]])
  ) +
  theme(legend.position = "bottom",
        panel.background = element_rect(color = "black", fill = NA)) 
```

```{r ferm-ri-npip}
plt <- ferm_interacting |> 
  group_by(i) |> 
  summarise(n_pip = mean(n_pip)) |> 
  left_join(sequence) |> 
  left_join(aa, by = c('res' = '1')) |>
  update(1, charge, "Positive") |>
  ggplot(aes(x = i, y = n_pip)) + 
  annotate(geom = "rect",
           xmin = 140, xmax = 164,
           ymin = 0, ymax = Inf,
           color = NA,
           fill = "black",
           alpha = 0.05) +
  geom_col(aes(fill = charge),
           alpha = 1,
           width = 0.8) +
  geom_text(aes(label = res, color = charge),
            fontface = "bold",
            vjust = -0.2,
            data = ~ filter(.x, n_pip > 0.1), show.legend = FALSE) +
  scale_fill_manual(
    values = CHARGE_COLORS, aesthetics = c("fill", "color"),
    name = "Net Charge at pH 7.4"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(breaks = seq(0, 400, 50)) +
  labs(y = expression(n[PIP[2]])) +
  coord_cartesian(clip = "off", xlim = c(0, 400)) +
  labs(
    x = "residue index",
    y = expression(bar(n)[PIP[2]])
    ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = c(0, 1),
    legend.justification = c(0, 1),
    legend.title = element_text(size = 12)
  )

grobs <- ggplotGrob(plt)

plt <- plt +
guides(
    color = guide_colorbar(
      barheight = unit(0.7, "grobheight", data = grobs$grobs),
      frame.colour = "black"
    )
)

anot <- ggplot(ferm_annotation) +
  aes(
    ymin = 0, ymax = 10,
    xmin = imin, xmax = imax,
    fill = domain) +
  geom_rect(show.legend = FALSE) +
  geom_vline(data = ferm_pip_sites, aes(xintercept = ri), color = "red") +
  geom_text(aes(y = 5, x = (imin + imax) / 2, label = domain),
            color = "black") +
  theme_void() +
  coord_cartesian(clip = "off", xlim = c(0, 400)) +
  scale_fill_discrete(type = ferm_colors)


(plt / anot) +
  plot_layout(heights = c(20, 1)) +
  plot_annotation()
```

## SI

### RMSD

```{r}
#| label: ferm-rmsf
ferm_rmsf |>
  ggplot(aes(residue, rmsd)) +
  geom_line() +
  annotate(geom = "rect",
    xmin = 135, xmax = 168,
    ymin = 0, ymax = Inf,
    color = NA,
    fill = "black",
    alpha = 0.05) +
  geom_rect(
    data = ferm_annotation,
    inherit.aes = FALSE,
    aes(xmin = imin, xmax = imax, fill = domain),
    ymin = 0, ymax = 0.03
  ) +
  labs(
    x = "residue index",
    y = "rmsf [nm]"
  ) +
  scale_fill_discrete(type = ferm_colors) +
  guides(fill = "none")
```


